---
title: 'Market Basket Analysis'
author: "Ashley Fidler"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Market Basket Analysis

```{r}
library("arules")
library("arulesViz")
library("tidyverse")
library("readxl")
library("knitr")
library("lubridate")
library("plyr")
```

```{r}
#Pre-Processing

retail <- read_excel("Online Retail (1).xlsx")          #loading in the data

retail <- retail[complete.cases(retail), ]          #filtering for only complete rows

retail %>% mutate(Description = as.factor(Description))     #Description data becomes a factor

retail %>% mutate(Country = as.factor(Country))             #Country data becomes as factor

retail$Date <- as.Date(retail$InvoiceDate)          #Character data becomes a date in a new variable 

TransTime <- format(retail$InvoiceDate, "%H:%M:%S")         #Separates time as a new variable

InvoiceNo <- as.numeric(as.character(retail$InvoiceNo))     #InvoiceNo data becomes numeric 

cbind(retail, TransTime)                            #Binds TransTime column into retail dataframe

cbind(retail, InvoiceNo)                            #Binds InvoiceNo column into retail dataframe

glimpse(retail)
```

```{r}
#Converting Data to Transaction Data 

transactionData <- ddply(retail, c("InvoiceNo", "Date"),
                         function(df1)paste(df1$Description, collapse = ","))           #Combines all info from an invoice number into 1 row 

transactionData$InvoiceNo <- NULL               #InvoiceNo will not be of use 

transactionData$Date <- NULL                    #Date will not be of use

colnames(transactionData) <- c("items")         #Rename column to items 

transactionData

write.csv(transactionData, "market_basket_transactions.csv", quote = FALSE, row.names = FALSE)
```

```{r}
#Transaction data loads into transaction class 

tr <- read.transactions("market_basket_transactions.csv", format = "basket", sep = ",")

tr

summary(tr)
```
There are 68 transactions (rows) and 546 items (columns) -- 546 is the product descriptions involved in the dataset, and 68 transactions are collections of these items.

Density is the percentage of non-zero cells in a sparse matrix (a sparse matrix is a matrix where most of the elements are zero). You can say it as the total number of items that are purchased divided by a possible number of items in that matrix
Density  = 68 * 546 * .02265137
Density = 841.000065

Element length distribution tells you how many transactions there are for 1-itemset, 2-itemset, 3-itemset, etc. The first row if the number of items and the second row is the number of transactions.

```{r}
#Item Frequency Plot 

if (!require("RColorBrewer")) {
    install.packages("RColorBrewer")
    library(RColorBrewer)
}
itemFrequencyPlot(tr, topN = 20, type = "absolute", col = brewer.pal(8, "Pastel2"), main = "Absolute Item Frequency Plot")      #type = "absolute" shows numeric frequencies of each item independently

itemFrequencyPlot(tr, topN = 20, type = "relative", col = brewer.pal(8, "Pastel2"), main = "Absolute Item Frequency Plot")      #type = "relative" shows how many times these items appear when compared to others
```

```{r}
#Generating Rules with APRIORI Algorithm

association.rules <- apriori(tr, parameter = list(supp = .05, conf = .8, maxlen = 10))         #confidence is .8, min support is .001 

summary(association.rules)
```

Parameter Specification: min_sup = .001 and min_confidence = .8 values with 3 items as max of items in a rule

Total number of rules: the set of 525376 rules

Distribution of Rule Length: a length of 3 items has the most rules (at 496798). a length of 2 items has the lowest number of rules (at 13623)

Summary of quality measures: min and max values for support, confidence, and lift 

Information used for creating rules: the data, support, and confidence we provided to the algorithm 

```{r}
#Looking at the top 10 rules 

inspect(head(sort(association.rules, by = "confidence"), 10))
```


```{r}
#Limiting number and size of rules 

shorter.association.rules <- apriori(tr, parameter = list(supp = .05, conf = .9, maxlen = 3))

summary(shorter.association.rules)

inspect(head(sort(shorter.association.rules, by = "confidence"), 10))
```

Parameter Specification: min_sup = .001 and min_confidence = .9 values with 3 items as max of items in a rule

Total number of rules: the set of 510155 rules

Distribution of Rule Length: a length of 3 items has the most rules (at 496599). a length of 2 items has the lowest number of rules (at 13556)


```{r}
#Finding rules related to given items 

aged.association.rules <- apriori(tr, parameter = list(supp = .001, conf = 0.8, maxlen = 3), appearance = list(default = "lhs", rhs = "HAND WARMER UNION JACK"))    #what do customers buy before buying "HAND WARMER UNION JACK"

inspect(head(sort(aged.association.rules, by = "confidence"), 5))
```

```{r}
#Customers who bought HAND WARMER UNION JACK also bought...

aged.association.rules <- apriori(tr, parameter = list(supp = .001, conf = 0.1, maxlen = 3), appearance = list(lhs = "HAND WARMER UNION JACK", default = "rhs"))     #keep HAND WARMER UNION JACK on lhs

inspect(head(sort(aged.association.rules, by = "confidence"), 5))
```















